Revisiting the admissibility of non-contextual hidden variable models in quantum mechanics

### 2018-11-14 version3supportFiles/PLA/non_contextual_pla_revised.tex
Added Kishor's changes and now working on changing the format to PLA

\documentclass[doublecol,british]{epl2} 
% or \documentclass[page-classic]{epl2} for one column style
\usepackage[normalem]{ulem}

%\setcounter{secnumdepth}{3}     
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{refstyle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{braket}
\geometry{verbose,tmargin=3cm,bmargin=2cm,lmargin=2cm,rmargin=2cm,headheight=2cm,headsep=2cm,footskip=1cm}
\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\lemref[1]{\ref{lem:#1}}}
\AtBeginDocument{\providecommand\propref[1]{\ref{prop:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}
\AtBeginDocument{\providecommand\defnref[1]{\ref{defn:#1}}}
\AtBeginDocument{\providecommand\tblref[1]{\ref{tbl:#1}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{paralist}
\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{xcolor}
%\usepackage{url}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\makeatother

\usepackage{babel}
  \providecommand{\definitionname}{Definition}
  \providecommand{\examplename}{Example}
  \providecommand{\lemmaname}{Lemma}
  \providecommand{\propositionname}{Proposition}
  \providecommand{\remarkname}{Remark}
  \providecommand{\theoremname}{Theorem}
  \providecommand{\notationname}{Notation}
\setlength{\itemindent}{0pt}

% \usepackage{comment}
% \excludecomment{sout}
%\let\endsout\relax

%%%%%%%%%%%%%%%%%%%

\title{Revisiting the admissibility of non-contextual hidden variable models in quantum mechanics}
% \title{Relexation of functional-consistency allows
% a non-contextual hidden variable model for quantum mechanics
%\\
% \sout{A non-contextual Hidden Variable Model for Quantum
% Mechanics}}
\shorttitle{Admissibility of non-contextual HV models in QM} 
%Non-contextual HV Model for QM} %Insert here a short version of the title if it exceeds 70 characters

\author{Atul Singh Arora\inst{1,2} \and Kishor Bharti\inst{3} \and Arvind\inst{1}}
\shortauthor{A. S. Arora \etal}

\institute{                    
  \inst{1} Indian Institute of Science Education and Research (IISER), Mohali,\\Sector 81 SAS Nagar 140306 Punjab India.\\
  \inst{2} Centre for Quantum Information and Communication, Ecole polytechnique de Bruxelles, CP 165,\\Universit\'e libre de Bruxelles (ULB), 1050 Brussels, Belgium \\
  %\inst{2} Universit\'e libre de Bruxelles, \\50 av. F. D. Roosevelt - CP 165/59, B-1050 Brussels, Belgium.\\
  \inst{3} Centre for Quantum Technologies (CQT), \\National University of Singapore (NUS) -  Block S15, 3 Science Drive 2, Singapore 117543.
}

\pacs{03.65.Ta}{Foundations of quantum mechanics}
%\pacs{}{}
% \pacs{nn.mm.xx}{First pacs description}
% \pacs{nn.mm.xx}{Second pacs description}
% \pacs{nn.mm.xx}{Third pacs description}

\abstract{ %
We construct a non-contextual
hidden variable model consistent with all { the}
kinematic predictions of quantum mechanics (QM). The famous
Bell-KS theorem shows that non-contextual models which
satisfy a further reasonable restriction are inconsistent
with QM. { In our construction,} we define a
weaker variant of this restriction which captures its
essence while still allowing a non-contextual description of
QM. This is in contrast to { the} contextual
 hidden variable {toy} models, such as
the one by Bell, { and}  brings out an interesting
alternate way of looking at QM. { The results}
also relate to { the Bohmian}
model, where it is harder to pin
down such features.}
\begin{document}

\maketitle

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.98\columnwidth]{block2}
\end{center}
\caption{Non-contextuality is not inconsistent
with kinematic predictions of QM} \label{fig:block}\end{figure}
{
The KS theorem is applicable to non-contextual HV models
which additionally satisfy \emph{functional-consistency---algebraic 
constraints obeyed by commuting quantum
observables must be satisfied by the HV model at the level
of individual outcomes}. 
One of the main justifications for imposing
this was the requirement that
{ the} observable $\hat A^2$ must depend on $\hat
A$ even at the HV level. This sounds reasonable because
otherwise one can construct HV models where these
observables { can} get mapped to random variables
with no relation to each other (see~\cite{KochenSpecker}). 
As we
will show, one can weaken this requirement into what we call
\emph{weak functional-consistency---functional-consistency is
demanded only when the observables in question have sharp
values}. This entails that even in the HV model the
observables will depend on each other where they should for
{ consistency} and not otherwise,
thereby capturing the essential idea without undue
constraints. The consequence of
demanding only the weaker variant is that the KS theorem no
longer applies and non-contextual HV models can be
consistently constructed. This is in stark contrast with
contextual HV models, such as the toy
model\footnote{introduced in connection with Gleason's
Theorem} by Bell~\cite{BellOnHiddenVariables} because
here we consider models where the algebraic constaints of 
QM are not obeyed in general by the HV model at the level of
individual outcomes. We demonstrate this contrast
by re-examining a `proof of contextuality' using
our model.}
% {
% A closer scrutiny of the arguments
% leading to the dichotomy between QM and HV models
% reveals that the algebraic constraints obeyed by
% quantum observables may not necessarily be obeyed
% by the predictions of an HV model at the level of
% individual outcomes and  that contextuality is 
% not a necessary feature of QM. Those HV
% models that obey such algebraic constraints
% possess a special property which we call
% ``multiplicativity''.  Enlarging the scope of HV
% models to include non-multiplicative models,
% permits us to construct an explicit non-contextual
% HV model consistent with 
% kinematic predictions of
% QM  %(see \figref{block}).
% The re-examination of the very proof of
% ``contextuality''  is used to illustrate
% non-multiplicativity, an intrinsic quantum
% feature, which we propose as a possible
% alternative to contextuality. 
% Non-mulitiplicativity thus might turn out to be
% the key quantum feature distinguishing the quantum
% world from the classical one and may have a role
% to play in QIP protocols. }
This provides a new
way of looking at the classical-quantum divide and at
the foundations of quantum mechanics.

@@

\begin{defn}
A prediction map $M$ is \emph{functionally-consistent} if
\begin{align*} &
M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi])
= \\ & \tilde
f(M(\hat{B}_{1},[\psi]),M(\hat{B}_{2},[\psi]),\dots
M(\hat{B}_{N},[\psi])), \end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables and
$[\psi]\in[\mathcal{H}]$. A
\emph{non functionally-consistent} map is one that is not
functionally-consistent.  
\end{defn} 
Note that if $M$ is taken to represent the
measurement outcome (in QM), then for states of
the system which are simultaneous eigenkets of
$\hat{B}_{i}$s, $M$ must clearly be
\emph{functionally-consistent}. It is, however, not obvious that
this property must always hold. For example,
consider two spin-half particles in the state 
$\left|1\right\rangle \otimes\left|1\right\rangle
$ written in the computational basis and 
operators 
$\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$
and
$\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$
written in terms of Pauli operators.
We must have $M(\hat{C})=-1$ while
$M(\hat{B}_{1})=\pm1$ and $M(\hat{B}_{2})=\pm1$
independently, according to QM, with probability
half. Here \emph{functional-consistency} clearly is not
required to hold { and indeed this is why it was demanded in addition to being consistent with QM by KS}.
Antithetically, it is clear that if one first
measures $\hat{B}_{1}$ and subsequently measures
$\hat{B}_{2}$, then the product of the results
must be $-1$. This is consistent with measuring
$\hat{C}$. { In our treatment, instead of imposing \emph{functional-consistency}, we demand its aforesaid weaker variant. It captures the same idea, however, only when it has a precise meaning according to QM.} { To this end we define}
\emph{weak functional-consistency} as follows.
\begin{defn} 
A prediction map $m$ has
\emph{weak functional-consistency} for a given
sequence map $s$, if \begin{align*}
&M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde
f(M(\hat{B}_{1},[\psi_{k_{1}}]),M(\hat{B}_{2},[\psi_{k_{2}}]),\dots,M(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*} 
where
$\{k_{1},k_{2},\dots,k_{N}\} \in\{
N! {\rm ~permutations~of~}  k'{\rm s} \}$, 
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables,
$[\psi_{i}]\in[\mathcal{H}]$ and
$[\psi_{k+1}] :=
S(\hat{B}_{k},[\psi_{k}],M(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.  
\label{defn:seqnMltpl}
\end{defn}

@@

\begin{proof} Due to Peres Mermin for
$\left|\mathcal{H}\right|\ge4$
\cite{Peres,Mermin}:
For a system composed of two spin-half particles 
consider the following set of operators 
\begin{equation}\label{eqn:PM}
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\mathbb{I}} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} &
\hat{\mathbb{I}}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right]
\end{equation}
which have the property that
all operators along a row (or column) commute. Further,
the product of rows (or columns) yields
$\hat{R}_{i}=\hat{\mathbb{I}}$ and
$\hat{C}_{j}=\hat{\mathbb{I}}\,(j\neq3)$,
$\hat{C}_{3}=-\hat{\mathbb{I}}$, ($\forall\,i,j$) where
$\hat{R}_{i}:=\prod_{j}\hat{A}_{ij}$,
$\hat{C}_{j}:=\prod_{i}\hat{A}_{ij}$. Let us
assume that $M$ exists. From property (b) of the
map, to get $M(\hat{C}_{3})=-1$ (as required by
property (a)), we must have an odd number of $-1$
assignments in the third column. In the remaining
columns, the number of $-1$ assignments must be
even (for each column). Thus, in the entire
square, the number of $-1$ assignments must be
odd. Let us use the same reasoning, but along the
rows. Since each $M(\hat{R}_{i})=1$, we must have
even number of $-1$ assignments along each row.
Thus, in the entire square, the number of $-1$
assignments must be even. We have arrived at a
contradiction and therefore we conclude 
that $M$ does not exist.
\end{proof}

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{functionally-consistent} must be incompatible with
QM. 
This leaves open an interesting possibility that non-contextual maps which have \emph{weak functional-consistency} could be consistent with QM.
{  Before proceeding to do so explicitly we observe that \emph{weak functional-consistency} is, in fact, a consequence of QM.}
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{weak functional consistency} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined in \defnref{seqnMltpl}.
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{weak functional-consistency}
must always hold.

\section{Construction}
We are now ready to describe our 
model explicitly.
Let the state of a finite dimensional quantum
system be $\left|\chi\right\rangle $. We wish to
assign a value to an arbitrary observable, given in terms of its spectral decomposition as,

>>>>

\documentclass[5p]{elsarticle}
%\documentclass[preprint,12pt]{elsarticle}
\usepackage{color}
%\tolerance=500
\raggedbottom
%\hyphenpenalty=100
%\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsthm}
\journal{Physics Letters A}
\bibliographystyle{elsarticle-num}
\makeatletter
\let\fnref\relax
\usepackage{refstyle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\lemref[1]{\ref{lem:#1}}}
\AtBeginDocument{\providecommand\propref[1]{\ref{prop:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}
\AtBeginDocument{\providecommand\defnref[1]{\ref{defn:#1}}}
\AtBeginDocument{\providecommand\tblref[1]{\ref{tbl:#1}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{paralist}
\usepackage{hyperref}
%\makeatother
\usepackage[english]{babel}
  \providecommand{\definitionname}{Definition}
  \providecommand{\examplename}{Example}
  \providecommand{\lemmaname}{Lemma}
  \providecommand{\propositionname}{Proposition}
  \providecommand{\remarkname}{Remark}
  \providecommand{\theoremname}{Theorem}
  \providecommand{\notationname}{Notation}
\begin{document}
\title{Revisiting the admissibility of non-contextual hidden 
variable models in quantum mechanics}
%\shorttitle{Admissibility of non-contextual HV models in QM} 
%\keywords{Non-contextuality, HV Model for QM}
\author{Atul Singh Arora}
\ead{atul.singh.arora@ulb.ac.be}
\address{Department of Physical Sciences,
Indian Institute of Science Education \&
Research (IISER) Mohali, Sector 81 SAS Nagar,
Manauli PO 140306 Punjab India.}
\address{Centre for Quantum Information and Communication,
Ecole polytechnique de Bruxelles, CP 165,\\Universit\'e
libre de Bruxelles (ULB), 1050 Brussels, Belgium}
\author{Kishor Bharti}
\ead{kishor.bharti1@gmail.com}
\address{Centre for Quantum Technologies (CQT), \\National University
of Singapore (NUS) -  Block S15, 3 Science Drive 2,
Singapore 117543.  } 
\author{Arvind}
\ead{arvind@iisermohali.ac.in}
\address{Department of Physical Sciences,
Indian Institute of Science Education \&
Research (IISER) Mohali, Sector 81 SAS Nagar,
Manauli PO 140306 Punjab India.}

%\shortauthor{A. S. Arora \etal}
%\pacs{03.65.Ta}{Foundations of quantum mechanics}
\begin{abstract}
We construct a non-contextual hidden variable model
consistent with all the kinematic predictions of quantum
mechanics (QM). The famous Bell-KS theorem shows that
non-contextual models which satisfy a further reasonable
restriction are inconsistent with QM. In our construction,
we define a weaker variant of this restriction which
captures its essence while still allowing a non-contextual
description of QM. This is in contrast to the contextual
hidden variable toy models, such as the one by Bell, and
brings out an interesting alternate way of looking at QM.
The results also relate to the Bohmian model, where it is
harder to pin down such features.
\end{abstract}
\maketitle

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.98\columnwidth]{block2}
\end{center}
\caption{Non-contextuality is not inconsistent
with kinematic predictions of QM} \label{fig:block}\end{figure}
{
The KS theorem is applicable to non-contextual HV models
which additionally satisfy \emph{functional-consistency ---
algebraic constraints obeyed by commuting quantum
observables must be satisfied by the HV model at the level
of individual outcomes}. 
One of the main justifications for imposing
this was the requirement that
{ the} observable $\hat A^2$ must depend on $\hat
A$ even at the HV level. This sounds reasonable because
otherwise one can construct HV models where these
observables { can} get mapped to random variables
with no relation to each other (see~\cite{KochenSpecker}). 
As we
will show, one can weaken this requirement into what we call
\emph{weak functional-consistency -- functional-consistency is
demanded only when the observables in question have sharp
values}. This entails that even in the HV model the
observables will depend on each other where they should for
{ consistency} and not otherwise,
thereby capturing the essential idea { without undue
constraints}. The consequence of
demanding only the weaker variant is that the KS theorem no
longer applies and non-contextual HV models can be
consistently constructed. This is in stark contrast with
contextual HV models, such as the toy
model\footnote{introduced in connection with Gleason's
Theorem} by Bell~\cite{BellOnHiddenVariables} because
here we consider models where the algebraic constaints of 
QM are not obeyed in general by the HV model at the level of
individual outcomes. We demonstrate this contrast
by re-examining a `proof of contextuality' using
our model.}
% {
% A closer scrutiny of the arguments
% leading to the dichotomy between QM and HV models
% reveals that the algebraic constraints obeyed by
% quantum observables may not necessarily be obeyed
% by the predictions of an HV model at the level of
% individual outcomes and  that contextuality is 
% not a necessary feature of QM. Those HV
% models that obey such algebraic constraints
% possess a special property which we call
% ``multiplicativity''.  Enlarging the scope of HV
% models to include non-multiplicative models,
% permits us to construct an explicit non-contextual
% HV model consistent with 
% kinematic predictions of
% QM  %(see \figref{block}).
% The re-examination of the very proof of
% ``contextuality''  is used to illustrate
% non-multiplicativity, an intrinsic quantum
% feature, which we propose as a possible
% alternative to contextuality. 
% Non-mulitiplicativity thus might turn out to be
% the key quantum feature distinguishing the quantum
% world from the classical one and may have a role
% to play in QIP protocols. }
This provides a new
way of looking at the classical-quantum divide and at
the foundations of quantum mechanics.

@@

\begin{defn}
A prediction map $M$ is \emph{functionally-consistent} iff
\begin{align*} &
M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi])
= \\ & \tilde
f(M(\hat{B}_{1},[\psi]),M(\hat{B}_{2},[\psi]),\dots
M(\hat{B}_{N},[\psi])), \end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables and
$[\psi]\in[\mathcal{H}]$. A
\emph{non functionally-consistent} map is one that is not
functionally-consistent.  
\end{defn} 
Note that if $M$ is taken to represent the
measurement outcome (in QM), then for states of
the system which are simultaneous eigenkets of
$\hat{B}_{i}$s, $M$ must clearly be
\emph{functionally-consistent}. It is, however, not obvious that
this property must always hold. For example,
consider two spin-half particles in the state 
$\left|1\right\rangle \otimes\left|1\right\rangle
$ written in the computational basis and 
operators 
$\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$
and
$\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$
written in terms of Pauli operators.
We must have $M(\hat{C})=-1$ while
$M(\hat{B}_{1})=\pm1$ and $M(\hat{B}_{2})=\pm1$
independently, according to QM, with probability
half. Here \emph{functional-consistency} clearly is not
required to hold { and indeed this is why it was demanded in addition to being consistent with QM by KS}.
Antithetically, it is clear that if one first
measures $\hat{B}_{1}$ and subsequently measures
$\hat{B}_{2}$, then the product of the results
must be $-1$. This is consistent with measuring
$\hat{C}$. { In our treatment, instead of imposing \emph{functional-consistency}, we demand its aforesaid weaker variant. It captures the same idea, however, only when it has a precise meaning according to QM.} { To this end we define}
\emph{weak functional-consistency} as follows.
\begin{defn} 
A prediction map $m$ has
\emph{weak functional-consistency} for a given
sequence map $s$, iff \begin{align*}
&M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde
f(M(\hat{B}_{1},[\psi_{k_{1}}]),M(\hat{B}_{2},[\psi_{k_{2}}]),\dots,M(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*} 
where
$\{k_{1},k_{2},\dots,k_{N}\} \in\{
N! {\rm ~permutations~of~}  k'{\rm s} \}$, 
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables,
$[\psi_{i}]\in[\mathcal{H}]$ and
$[\psi_{k+1}] :=
S(\hat{B}_{k},[\psi_{k}],M(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.  
\label{defn:seqnMltpl}
\end{defn}

@@

\begin{proof} Peres Mermin [PM]
($\left|\mathcal{H}\right|\ge4$)
\cite{Peres,Mermin}:
For a system composed of two spin-half particles 
consider the following set of operators 
\begin{equation*}
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\mathbb{I}} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} &
\hat{\mathbb{I}}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right]
\end{equation*}
which have the property that
all operators along a row (or column) commute. Further,
the product of rows (or columns) yields
$\hat{R}_{i}=\hat{\mathbb{I}}$ and
$\hat{C}_{j}=\hat{\mathbb{I}}\,(j\neq3)$,
$\hat{C}_{3}=-\hat{\mathbb{I}}$, ($\forall\,i,j$) where
$\hat{R}_{i}:=\prod_{j}\hat{A}_{ij}$,
$\hat{C}_{j}:=\prod_{i}\hat{A}_{ij}$. Let us
assume that $M$ exists. From property (b) of the
map, to get $M(\hat{C}_{3})=-1$ (as required by
property (a)), we must have an odd number of $-1$
assignments in the third column. In the remaining
columns, the number of $-1$ assignments must be
even (for each column). Thus, in the entire
square, the number of $-1$ assignments must be
odd. Let us use the same reasoning, but along the
rows. Since each $M(\hat{R}_{i})=1$, we must have
even number of $-1$ assignments along each row.
Thus, in the entire square, the number of $-1$
assignments must be even. We have arrived at a
contradiction and therefore we conclude 
that $M$ does not exist
\end{proof}

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{functionally-consistent} must be incompatible with
QM. 
This leaves open an interesting possibility that non-contextual maps which have \emph{weak functional-consistency} could be consistent with QM.
{  Before proceeding to do so explicitly we observe that \emph{weak functional-consistency} is, in fact, a consequence of QM.}
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{weak functional consistency} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined (in \defnref{seqnMltpl})
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{weak functional-consistency}
must always hold.

\section{Construction}
We are now ready to describe our 
model explicitly.
Let the state of a finite dimensional quantum
system be $\left|\chi\right\rangle $. We wish to
assign a value to an arbitrary observable
$\hat{A}=\sum_{a}a\left|a\right\rangle
\left\langle a\right|$, which has eigenvectors $\{
\vert a_j\rangle \}$. The corresponding ordered
eigenvalues are $\{a_j\}$ such that $a_{\rm
min}=a_1$ and $a_{\rm max}=a_n$.  Our HV model for
QM assigns values in the following three steps:
\setdefaultleftmargin{0pt}{}{}{}{}{}
\begin{enumerate}
\item
{\bf  Initial HV:} Pick a number
$c\in[0,1]$, from a uniform random distribution.\
\item
{\bf
Assignment or Prediction:}
 The value assigned to
$\hat{A}$ is given by finding the smallest $a$
s.t.  $c\le\sum_{a'=a_{\text{min}}}^{a}\left|\left\langle
a'|\chi\right\rangle \right|^{2},$ viz. we have
specified a prediction map, $M(\hat{A},[\chi ])=a$.
\item
{\bf  Update:} After measuring an operator, the state is
updated (collapsed) in accordance with the rules
of QM. This completely specifies the sequence map
$S$.
\end{enumerate}
The above HV model works for all quantum systems,
however, to illustrate its working,
consider the example of a spin-half particle in the state
$\left|\chi\right\rangle
=\cos\theta\left|0\right\rangle
+\sin\theta\left|1\right\rangle $ and the observable
$\hat{A}=\hat{\sigma}_{z}=\left|0\right\rangle
\left\langle 0\right|-\left|1\right\rangle
\left\langle 1\right|$. Now, according to the
postulates of this theory, $M(\hat{A},[\chi])=+1$
if
the randomly generated value for 
$c\le\cos^{2}\theta$ else $\hat{A}$ is assigned
$-1$; it then follows,  from $c$ being uniformly
random in $[0,1]$ that the statistics agree with
the predictions of QM {\it i.e.}  the
Born rule. 
The assignment described by the
prediction map $M$ is non-contextual since, given
an operator and a state (alongwith the HV $c$), the value is
uniquely assigned. The map $M$ is, however,
\emph{non functionally-consistent}.
